Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-04 intro.tex
add sentence to intro

To answer this question and for the purposes of this paper we focus on real-valued (possibly) noisy scalar time series which appear \emph{complicated}, i.e., they exhibit interesting behavior, for instance, not of low period, strictly monotonic, or constant. We make no assumptions about the generating process's properties,(e.g., linear, nonlinear, deterministic, stochastic, stationary, non-stationary, etc.) and we attempt to measure empirically where on the previously mentioned spectrum of \emph{complexity} the time series exists.
We argue that an extension of \emph{permutation entropy}
\cite{bandt2002per}---a method for approximating the entropy of a real-valued-finite-length-noisy time series through ordinal analysis---is an effective way to explore that conjecture. In light of this, for the purposes of this paper we define \emph{complexity} as an approximation of Shannon \emph{entropy}, which we define and explain rigorously in Section \ref{sec:meaComplex}. In this fashion a random walk time series (which exhibits high entropy) is purely complex, where as a periodic signal (which exhibits low entorpy) has low complexity. {\color{blue} Ryan: Insert your sentence here, Something like We are not the first people to attempt to quantify complexity of time series with entropy(citations) However, these methods rely on binning which is bad because of biasing (citations) which we explain further in Section \ref{sec:related}.}

>>>>

To answer this question and for the purposes of this paper we focus on real-valued (possibly) noisy scalar time series which appear \emph{complicated}, i.e., they exhibit interesting behavior, for instance, not of low period, strictly monotonic, or constant. We make no assumptions about the generating process's properties,(e.g., linear, nonlinear, deterministic, stochastic, stationary, non-stationary, etc.) and we attempt to measure empirically where on the previously mentioned spectrum of \emph{complexity} the time series exists.
We argue that an extension of \emph{permutation entropy}
\cite{bandt2002per}---a method for approximating the entropy of a real-valued-finite-length-noisy time series through ordinal analysis---is an effective way to explore that conjecture. In light of this, for the purposes of this paper we define \emph{complexity} as an approximation of Shannon \emph{entropy}, which we define and explain rigorously in Section \ref{sec:meaComplex}. In this fashion a random walk time series (which exhibits high entropy) is purely complex, where as a periodic signal (which exhibits low entorpy) has low complexity. The permutation entropy is ideal in a situation like this as it converges on the true entropy value, unlike other partitioning techniques used in the past which either require specific knowledge of the generating process or produce biased values of the entropy~\cite{bollt2001}.
